---
title: "hw6_p8105"
output: github_document
date: "2024-11-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(ggridges)
library(dplyr)
set.seed(1)
```

## Problem 2 

**2A. Data export and cleaning**
```{r homicide dataset, warning=FALSE}
homicide = read.csv("./data/homicide-data.csv") %>% 
  unite(city_state, c(city, state), sep = ", ") %>% 
  mutate(
    disposition = str_replace(disposition, "Closed by arrest", "Solved"), 
    disposition = str_replace(disposition, "Closed without arrest", "Unsolved"), 
    disposition = str_replace(disposition, "Open/No arrest", "Unsolved"), 
    disposition_binary = ifelse(disposition == "Solved", 1, 0),
    victim_age = as.numeric(victim_age)
  ) %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) %>%
  filter(victim_race %in% c("White", "Black")) %>% 
  filter(victim_sex %in% c("Female", "Male")) %>% 
  select(city_state, victim_race, victim_age, victim_sex, disposition_binary) %>% 
  drop_na()
    
```

**2B. Baltimore, MD**
```{r Baltimore, warning=FALSE, message=FALSE}
baltimore_glm = homicide %>% 
  filter(city_state == "Baltimore, MD") %>%  
  glm(disposition_binary ~ victim_age + victim_race + victim_sex, family = binomial(), data = .)

tidy_baltimore = broom::tidy(baltimore_glm)

CI = confint(baltimore_glm) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "term") %>%
  rename("conf_low" = "2.5 %", "conf_high" = "97.5 %")
  
baltimore_data = left_join(tidy_baltimore, CI, by = "term") %>% 
  mutate(
    OR = exp(estimate),
    conf_low2 = exp(conf_low), 
    conf_high2 = exp(conf_high)
    ) %>% 
  rename("log_OR" = "estimate") %>% 
  select(term, OR, conf_low2, conf_high2) %>% 
  filter(term == "victim_sexMale")
  
knitr::kable(baltimore_data)

#can't use knitr in piping
```

**2C. All cities**
```{r homicide all, warning=FALSE, message=FALSE}
homicide_function = function(df) {
  glm_result = glm(disposition_binary ~ victim_age + victim_race + victim_sex, family = binomial(), data = df)
}

homicide_glm = homicide %>% 
  group_by(city_state) %>% 
  nest() %>%   
  mutate(
    glm_result = map(data, homicide_function), 
    tidy_result = map(glm_result, broom::tidy), 
    confint_result = map(glm_result, ~ as.data.frame(confint(.x)))
  ) %>% 
  mutate(
    stat_result = map2(tidy_result, confint_result, ~ {
      tidy_df <- .x 
      ci <- .y 
      colnames(ci) <- c("conf_low", "conf_high") 
      ci$term <- rownames(ci)
      left_join(tidy_df, ci, by = "term")
    })
  )
  
#cannot put CI as part of the homicide function, as glm_data will end up being CI result instead of glm result
#do separate mutate functions

homicide_stat = homicide_glm %>% 
  select(city_state, stat_result) %>% 
  unnest(stat_result) %>% 
  mutate(
    OR = exp(estimate),
    conf_low2 = exp(conf_low), 
    conf_high2 = exp(conf_high),
    city_state = as.factor(city_state)
    ) %>% 
  rename("log_OR" = "estimate") %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, OR, conf_low2, conf_high2)

#knitr::kable(homicide_stat)
```

**2D. Plotting homicide stat**
```{r homicide plot}
homicide_stat = homicide_stat %>% 
  arrange(OR) %>% 
  mutate(city_state = fct_inorder(city_state))

ggplot(homicide_stat, aes(x = OR, y = city_state)) + 
    geom_point() +
    geom_errorbar(aes(xmin = conf_low2, xmax = conf_high2)) +
    labs(
    x = "Estimates",
    y = "City, State",
    title = "OR & CI for homicide case outcome based on victim's gender"
  ) +
  theme_minimal()

#need to exp the CI too otherwise the OR falls outside of the CI range (bc the CI was still the original CI log)
#fct_reorder not working
#ggplot still change my df into list
```

My `fct_reorder` and `arrange` function didn't work, so I couldn't organize the plot based on the OR. But we can see that in most city-states, male victim's cases have lower odds of being solved compared to female victim's. Only 3 city-states where male victim's cases have higher odds of being solved, which are Stockton (CA), Fresno (CA), and Albuquerque (NM)- all of which has wide confidence interval. 

## Problem 3

**3a. Load and clean the data for regression analysis**
```{r bwt data, message=FALSE, warning=FALSE}
bwt_masterdata = homicide = read.csv("./data/birthweight.csv") %>% 
  mutate(
    babysex = as.factor(babysex),
    malform = as.factor(malform),
    smoken = as.integer(round(smoken))
  ) %>% 
  drop_na()

#ggplot(birthweight_masterdata, aes(x=wtgain)) +
  #geom_histogram()
#fincome, momage, race,  parity, pnumlbw, pnumsga, smoken are all non-normal distribution
```

**3b. Propose a regression model for birthweight. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.**

```{r bwt model, warning=FALSE, message=FALSE, results='hide'}
#loading packages
library(modelr)

#model building
fit = lm(bwt ~ smoken + wtgain, data = bwt_masterdata) 

tidy_lm = fit%>% 
  broom::tidy()

#add residuals
modelr::add_residuals(bwt_masterdata, fit)

#add predictions
modelr::add_predictions(bwt_masterdata, fit)

#fit cannot be used for modelr if I already tidy them, so I created new df (tidy_lm) to tidy the data

#plot
bwt_model = bwt_masterdata %>% 
  modelr::add_residuals(fit) %>% 
  modelr::add_predictions(fit)

ggplot(bwt_model, aes(x = pred, y = resid)) +
  geom_point(alpha=0.5) +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Model Residuals vs. Fitted Values (Model 1)"
  ) +
  theme_minimal()
```

The model hypothesized that baby's birthweight is influenced by mother's weightgain and smoking habit during the pregnancy. There are strong evidence that smoking during pregnancy negatively affects birthweight. Meanwhile, while there are evidence of mother's weight gain positively affects birthweight, this effect apparently dependent on the timing of the weight gain, with mothers who gain weight during the 2nd trimester seems to positively affects birthweight the most. As the outcome is not binary, I used `lm` function with `bwt` as the outcome and `smoken` and `wtgain` as the independent predictors. I test the `modelr` package individually to ensure the code is correct, then merge them in `bwt_model` dataframe. Then I made the plot, with prediction in the x axis and residuals in the y axis. 

**3c. Compare your model** to two others:

- One using length at birth and gestational age as predictors (main effects only)
- One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

**Model 2**
```{r model 2}
#load packages
library(mgcv)

#length and GA as predictor
fit2 = lm(bwt ~ blength + gaweeks, data = bwt_masterdata)

tidy_lm2 = fit2%>% 
  broom::tidy()

bwt_model2 = bwt_masterdata %>% 
  modelr::add_residuals(fit2) %>% 
  modelr::add_predictions(fit2)

ggplot(bwt_model2, aes(x = pred, y = resid)) +
  geom_point(alpha=0.5) +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Model Residuals vs. Fitted Values (Model 2)"
  ) +
  theme_minimal()

#fit2 is much different from fit in distribution
```


**Model 3**
```{r model 3}
#head circumference, length, sex
fit3 = lm(bwt ~ bhead*blength*babysex + bhead + blength + babysex, data = bwt_masterdata)

tidy_lm3 = fit3%>% 
  broom::tidy()

bwt_model3 = bwt_masterdata %>% 
  modelr::add_residuals(fit3) %>% 
  modelr::add_predictions(fit3)

ggplot(bwt_model3, aes(x = pred, y = resid)) +
  geom_point(alpha=0.5) +
  labs(
    x = "Fitted Values",
    y = "Residuals",
    title = "Model Residuals vs. Fitted Values (Model 3)"
  ) +
  theme_minimal()

#fit3 also different in distributiion
```

**comparison**
```{r cross_validation}
cv_df1 = crossv_mc(bwt_masterdata, n = 100, test = 0.2) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

cv_df1 = cv_df1 %>% 
  mutate(
    fit_mod  = map(train, \(df) lm(bwt ~ smoken + wtgain, data = bwt_masterdata)),
    fit2_mod  = map(train, \(df) lm(bwt ~ blength + gaweeks, data = bwt_masterdata)),
    fit3_mod = map(train, \(df) lm(bwt ~ bhead*blength*babysex + bhead + blength + babysex, data = bwt_masterdata))
  ) %>% 
  mutate(
    rmse_fit = map2_dbl(fit_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_fit2 = map2_dbl(fit2_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_fit3 = map2_dbl(fit3_mod, test, \(mod, df) rmse(model = mod, data = df))
  )

cv_df1 %>%  
  select(starts_with("rmse")) %>%  
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>%  
  mutate(model = fct_inorder(model)) %>%  
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() + 
  labs(
    x = "Model name",
    y = "rmse",
    title = "Cross validation"
  )
```

As fit3 (model3) has the lowest RMSE, it has the best fit. 
