---
title: "hw6_p8105"
output: github_document
date: "2024-11-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(broom)
library(ggridges)
library(dplyr)
set.seed(1)
```

## Problem 2 

**2A. Data export and cleaning**
```{r homicide dataset, warning=FALSE}
homicide = read.csv("./data/homicide-data.csv") %>% 
  unite(city_state, c(city, state), sep = ", ") %>% 
  mutate(
    disposition = str_replace(disposition, "Closed by arrest", "Solved"), 
    disposition = str_replace(disposition, "Closed without arrest", "Unsolved"), 
    disposition = str_replace(disposition, "Open/No arrest", "Unsolved"), 
    disposition_binary = ifelse(disposition == "Solved", 1, 0),
    victim_age = as.numeric(victim_age)
  ) %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) %>%
  filter(victim_race %in% c("White", "Black")) %>% 
  filter(victim_sex %in% c("Female", "Male")) %>% 
  select(city_state, victim_race, victim_age, victim_sex, disposition_binary) %>% 
  drop_na()
    
```

**2B. Baltimore, MD**
```{r Baltimore, warning=FALSE, message=FALSE}
baltimore_glm = homicide %>% 
  filter(city_state == "Baltimore, MD") %>%  
  glm(disposition_binary ~ victim_age + victim_race + victim_sex, family = binomial(), data = .)

tidy_baltimore = broom::tidy(baltimore_glm)

CI = confint(baltimore_glm) %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "term") %>%
  rename("conf_low" = "2.5 %", "conf_high" = "97.5 %")
  
baltimore_data = left_join(tidy_baltimore, CI, by = "term") %>% 
  mutate(
    OR = exp(estimate),
    conf_low2 = exp(conf_low), 
    conf_high2 = exp(conf_high)
    ) %>% 
  rename("log_OR" = "estimate") %>% 
  select(term, OR, conf_low2, conf_high2) %>% 
  filter(term == "victim_sexMale")
  
knitr::kable(baltimore_data)

#can't use knitr in piping
```

**2C. All cities**
```{r homicide all, warning=FALSE, message=FALSE}
homicide_function = function(df) {
  glm_result = glm(disposition_binary ~ victim_age + victim_race + victim_sex, family = binomial(), data = df)
}

homicide_glm = homicide %>% 
  group_by(city_state) %>% 
  nest() %>%   
  mutate(
    glm_result = map(data, homicide_function), 
    tidy_result = map(glm_result, broom::tidy), 
    confint_result = map(glm_result, ~ as.data.frame(confint(.x)))
  ) %>% 
  mutate(
    stat_result = map2(tidy_result, confint_result, ~ {
      tidy_df <- .x 
      ci <- .y 
      colnames(ci) <- c("conf_low", "conf_high") 
      ci$term <- rownames(ci)
      left_join(tidy_df, ci, by = "term")
    })
  )
  
#cannot put CI as part of the homicide function, as glm_data will end up being CI result instead of glm result
#do separate mutate functions
#how to extract the confint result?

homicide_stat = homicide_glm %>% 
  select(city_state, stat_result) %>% 
  unnest(stat_result) %>% 
  mutate(
    OR = exp(estimate),
    conf_low2 = exp(conf_low), 
    conf_high2 = exp(conf_high),
    city_state = as.factor(city_state)
    ) %>% 
  rename("log_OR" = "estimate") %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, OR, conf_low2, conf_high2)

knitr::kable(homicide_stat)
```

**2D. Plotting homicide stat**
```{r homicide plot}
homicide_stat = homicide_stat %>% 
  arrange(OR) %>% 
  mutate(city_state = fct_reorder(city_state, OR))

ggplot(homicide_stat, aes(x = OR, y = city_state)) + 
    geom_point() +
    geom_errorbar(aes(xmin = conf_low2, xmax = conf_high2))

#need to exp the CI too otherwise the OR falls outside of the CI range (bc the CI was still the original CI log)
#fct_reorder not working
#ggplot still change my df into list
```

Comment on the plot

## Problem 3
3a. Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r bwt data, message=FALSE, warning=FALSE}
bwt_masterdata = homicide = read.csv("./data/birthweight.csv") %>% 
  mutate(
    babysex = as.factor(babysex),
    malform = as.factor(malform),
    smoken = as.integer(round(smoken))
  ) %>% 
  drop_na()

#ggplot(birthweight_masterdata, aes(x=wtgain)) +
  #geom_histogram()
#fincome, momage, race,  parity, pnumlbw, pnumsga, smoken are all non-normal distribution
```

3b. Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.
```{r bwt model}
#loading packages


#model building
fit = lm(bwt ~ smoken + wtgain, data = bwt_masterdata) %>% 
  broom::tidy()

#add residuals
modelr::add_residuals(bwt_masterdata, fit)

#Error in eval(response_var(model), as.data.frame(data)) : object 'term' not found

#add predictions
modelr::add_predictions(bwt_masterdata, fit)

#Error in UseMethod("predict") :   no applicable method for 'predict' applied to an object of class "c('tbl_df', 'tbl', 'data.frame')"
```


3c. Compare your model to two others:

- One using length at birth and gestational age as predictors (main effects only)
- One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.